Quantum Pseudo-Telepathy

### 2023-06-06 paper/main.tex
quantum sol example

As seen in Section~\ref{sec:classic-sol}, the \emph{magic} square is
called this way because, under classical assumptions, there is no such
a square: it is impossible to construct a \(3 \times 3\) matrix filled
with \(\pm 1\)'s, with the constraints that the product of every row
is \(+1\) and the product of every column is \(-1\). We can, however,
construct a ``quantum magic'' square. Figure~\ref{fig:quantum} shows
the \(3 \times 3\) grid that Alice and Bob can use to win the Magic
Square game systematically.  Notice that the eigenvalues of the
observables at each cell are \(\pm 1\). Furthermore, the product of
every row is \(+I\) and the product of every column is \(-I\). (The
original observables did not have the negative \(X\) and \(Z\) in the
last row, so these two properties are not mandatory~\cite{Mermin1990,
  aravind:xxx}.  We have adapted the observables merely to be
``consistent'' with the properties of the classical magic square, but
what really matters for Alice and Bob are the outcomes of the
measurements, which they will use to fill the entries asked by
Charlie.)

@@

\begin{tcbexample}{Quantum Solution}{quantum-sol}
  Suppose that Charlie assigns the first row to Alice. To decide the
  value of the first cell, she (based on the \(3 \times 3\) grid she
  and Bob have prepared) measures the observable \(I \otimes Z\)
  on her two qubits. The possible outcomes are
  \begin{equation}
    \label{eq:outcomes-1r}
    \operatorname{Out}((I \otimes Z)_{A} \otimes (I \otimes I)_{B}) =
    \begin{cases}
      +1, \text{ proj. onto } \{\ket{i0jk} \mid i, j, k \in \{0, 1\}\}
        \text{ with prob. } \frac{1}{2} \\
      -1, \text{ proj. onto } \{\ket{i1jk} \mid i, j, k \in \{0, 1\}\}
        \text{ with prob. } \frac{1}{2}, \\
    \end{cases}
  \end{equation}
  where the subscript \(A1\) denotes the state after Alice's first
  measurement. Suppose that Alice has observed \(+1\). Notice that
  only two of the components of the \(+1\)-eigenspace are seen in the
  initial state \(\ket{\psi}\): \(\ket{0000}\) and
  \(\ket{1010}\). Therefore, \(\ket{\psi}\) collapses to
  \begin{equation}
    \ket{\psi}_{A1} =  \frac{1}{\sqrt{2}}(\ket{0000} + \ket{1010}).
  \end{equation}
  For the second cell, Alice measures the observable \(Z \otimes I\),
  whose possible outcomes are
    \begin{equation}
    \label{eq:outcomes-1r}
    \operatorname{Out}((Z \otimes I)_{A} \otimes (I \otimes I)_{B}) =
    \begin{cases}
      +1, \text{ proj. onto } \{\ket{0ijk} \mid i, j, k \in \{0, 1\}\}
        \text{ with prob. } \frac{1}{2} \\
      -1, \text{ proj. onto } \{\ket{1ijk} \mid i, j, k \in \{0, 1\}\}
        \text{ with prob. } \frac{1}{2}. \\
    \end{cases}
  \end{equation}
  Suppose, now, that Alice has observed the outcome \(-1\). Notice that
  only one component of the \(-1\)-eigenspace is seen in the state
  \(\ket{\psi}_{A1}\). Hence, the state collapses to
  \begin{equation}
    \ket{\psi}_{A2} =  \ket{1010}.
  \end{equation}
\end{tcbexample}

\section{Quantum Unitaries (? research)}
\label{sec:quantum-unitaries}

>>>>

As seen in Section~\ref{sec:classic-sol}, the \emph{magic} square is
called this way because, under classical assumptions, there is no such
a square: it is impossible to construct a \(3 \times 3\) matrix filled
with \(\pm 1\)'s, with the constraints that the product of every row
is \(+1\) and the product of every column is \(-1\). We can, however,
construct a ``quantum magic'' square. Figure~\ref{fig:quantum} shows
the \(3 \times 3\) grid that Alice and Bob can use to win the Magic
Square game systematically.  Notice that the eigenvalues of the
observables at each cell are \(\pm 1\). Furthermore, the product of
every row is \(+I\) and the product of every column is \(-I\).

@@

\begin{tcbexample}{Quantum Solution}{quantum-sol}
  Suppose that Charlie assigns the first row to Alice. To decide the
  value of the first cell, she (based on the \(3 \times 3\) grid she
  and Bob have prepared) measures the observable \(I \otimes Z\)
  on her two qubits. The possible outcomes are
  \begin{equation}
    \operatorname{Out}((I \otimes Z)_{A} \otimes (I \otimes I)_{B}) =
    \begin{cases}
      +1, \text{ proj. onto } \{\ket{i0jk} \mid i, j, k \in \{0, 1\}\}
        \text{ with prob. } \frac{1}{2} \\
      -1, \text{ proj. onto } \{\ket{i1jk} \mid i, j, k \in \{0, 1\}\}
        \text{ with prob. } \frac{1}{2},
    \end{cases}
  \end{equation}
  Suppose that Alice has observed \(+1\). Notice that
  only two of the components of the \(+1\)-eigenspace are seen in the
  initial state \(\ket{\psi}\): \(\ket{0000}\) and
  \(\ket{1010}\). Therefore, \(\ket{\psi}\) collapses to
  \begin{equation}
    \ket{\psi}_{A1} =  \frac{1}{\sqrt{2}}(\ket{0000} + \ket{1010}),
  \end{equation}
  where the subscript \(A1\) denotes the state after Alice's first
  measurement.  For the second cell, Alice measures the observable
  \(Z \otimes I\), whose possible outcomes are
  \begin{equation}
    \operatorname{Out}((Z \otimes I)_{A} \otimes (I \otimes I)_{B}) =
    \begin{cases}
      +1, \text{ proj. onto } \{\ket{0ijk} \mid i, j, k \in \{0, 1\}\}
        \text{ with prob. } \frac{1}{2} \\
      -1, \text{ proj. onto } \{\ket{1ijk} \mid i, j, k \in \{0, 1\}\}
        \text{ with prob. } \frac{1}{2}.
    \end{cases}
  \end{equation}
  Suppose, now, that Alice has observed the outcome \(-1\). Notice that
  only one component of the \(-1\)-eigenspace is seen in the state
  \(\ket{\psi}_{A1}\). Hence, the state collapses to
  \begin{equation}
    \ket{\psi}_{A2} =  \ket{1010}.
  \end{equation}
  Finally, Alice measures the observable \(Z \otimes Z\), whose
  only possible outcomes is
  \begin{equation}
    \operatorname{Out}((Z \otimes Z)_{A} \otimes (I \otimes I)_{B}) =
    \begin{cases}
      -1, \text{ proj. onto } \{\ket{i\overline{i}jk} \mid i, j, k \in \{0, 1\}\}
        \text{ with prob. } 1,
    \end{cases}
  \end{equation}
  where \(\overline{i}\) is the \texttt{not} operator. Notice that,
  because \(\ket{\psi}_{A2}\) falls exactly into the
  \(-1\)-eigenspace, the only possible outcome is \(-1\), and the
  state does not change, i.e. \(\ket{\psi}_{A3} = \ket{\psi}_{A2}\).
  Alice, then, fills the first row with the values \(\{+1, -1, -1\}\),
  whose product is, as required, \(+1\). Upon receiving Alice's
  answer, Charlie goes to meet Bob. Suppose he assigns the second
  column to Bob. To fill in the value of the first cell (which is the
  intersection of the first row and second column), Bob measures the
  observable \(Z \otimes I\). Notice that \(\ket{\psi}_{A3}\) is
  entirely in the \(-1\)-eigenspace; thus, Bob will for sure observe
  the outcome \(-1\) ---~which matches Alice's answer~--- and the
  state will not change: \(\ket{\psi}_{B1} = \ket{\psi}_{A3}\). Moving
  on to the second cell, Bob measures the observable \(I \otimes X\),
  whose possible outcomes are (for readability, we are showing only
  the components of the eigenspaces that are not orthogonal to
  \(\ket{\psi}_{B1}\))
  \begin{equation}
    \operatorname{Out}((I \otimes I)_{A} \otimes (I \otimes X)_{B}) =
    \begin{cases}
      +1, \text{ proj. onto } \frac{1}{\sqrt{2}}(\ket{1010} + \ket{1011})
        \text{ with prob. } \frac{1}{2} \\
      -1, \text{ proj. onto } \frac{1}{\sqrt{2}}(\ket{1010} - \ket{1011})
        \text{ with prob. } \frac{1}{2}.
    \end{cases}
  \end{equation}
  Assume that Bob has observed the outcome \(+1\). Then, the state
  collapses to
  \begin{equation}
    \ket{\psi}_{B2} =  \frac{1}{\sqrt{2}}(\ket{1010} + \ket{1011}).
  \end{equation}
  Finally, Bob measures \(-Z \otimes X\) on his two qubits.
  The only possible outcome is
  \begin{equation}
    \operatorname{Out}((I \otimes I)_{A} \otimes (-Z \otimes X)_{B}) =
    \begin{cases}
      +1, \text{ proj. onto } \frac{1}{\sqrt{2}}(\ket{1010} + \ket{1011})
      \text{ with prob. } 1,
    \end{cases}
  \end{equation}
  and the state does not change, i.e.
  \(\ket{\psi}_{B3} = \ket{\psi}_{B2}\). Alice's answer satisfies the
  property of the rows (that the product of the entries is equal to
  \(+1\)) and Bob's answer satisfies the property of the columns
  (product is equal to \(-1\)).  Moreover, Alice and Bob agreed in the
  intersection of the row and the column.  Hence, they won this round!
\end{tcbexample}
\section{Quantum Unitaries (? research)}
\label{sec:quantum-unitaries}
